Chapter3_ideas

### 2019-12-17 results/Chapter3_ideas99.Rmd
working on finalizing Chap3 Outline

Data used in the paper
human data- Gruhn et al 2013 (plotted below, 755 cells)
Sun et al 2006
Oliver-Bonet et al 2007 
Lian et al 2008
Lu et al 2012
Hou et al 2013

@@

@@

hs.CO.length_plot <- ggplot(Melted.Hs.male_MLH1, aes(y=CO, x=Chrm.id))+geom_jitter(aes( color=row.name, alpha = .2))+ 
  ggtitle("Human male CO Counts")+theme(legend.position="none")
hs.CO.length_plot
```

@@

Heads of the cell-wide summary statistics

@@

                            nbivs = length(unique(Chrm.id)),
                            total.SC = sum(axis.length),
                            mean.SC = mean(axis.length),
                            cV.SC = cv(axis.length)
)
head(Hs.SC_cell_wide)

@@

@@

I didn't test the axis length trait yet.

@@

@@

                 totCO = sum(hand.foci.count),
                 total.SC = sum(SC.length),
                 mean_co = mean(hand.foci.count, na.rm = TRUE),
                 mean_sc = mean(SC.length),
                   var_co = var(hand.foci.count),
                   var_sc = var(SC.length),
                 cv_co.count = cv(hand.foci.count),
                 cv_SC.length = cv(SC.length)
)

@@

                 totCO = sum(hand.foci.count, na.rm = TRUE),
                  total.SC = sum(SC.length, na.rm = TRUE),
                 mean_co = mean(hand.foci.count, na.rm = TRUE),
                 mean_sc = mean(SC.length, na.rm = TRUE),
                   var_co = var(hand.foci.count),
                   var_sc = var(SC.length),
                 cv_co.count = cv(hand.foci.count),
                 cv_SC.length = cv(SC.length)
)

@@

>>>>

human Data used in the paper

Gruhn et al 2013 - T. Hassold,
4660 prophase sperm from 56 biopsies, 2000 prophase oocytes

Sun et al 2006 --  F. Sun, 824 MLH1 FISH spreads(~100 cells / 10 men)

Oliver-Bonet et al 2007 -- Oliver-
Bonet  (RPA foci for 5 different SCs ~50 cells, ~400 cells with FISH + MLH1). problem if these data are included in big dataset

Lian et al 2008 - S. Xie, (maybe similar data as sun et al 2008?)
Lian, Jie, et al. "Variation in crossover interference levels on individual chromosomes from human males." Human molecular genetics 17.17 (2008): 2583-2594.

Lu et al 2012 - 
Hou et al 2013 - no SC lengths, (CO sounts 68 cells (oocytes?))

Human data are from mulitple citations; Gruhn et al 2013, Sun et al 2006, Oliver-Bonet et al 2007, Lian et al 2008, Lu et al 2012, Hou et al 2013.

Cytological Studies of Human Meiosis: Sex-Specific
Differences in Recombination Originate at, or Prior to,
Establishment of Double-Strand Breaks

@@

Melted.Hs.male_axis$Chrm.id <- factor(Melted.Hs.male_axis$Chrm.id, ordered = TRUE, levels=c("chr1","chr2","chr3",  "chr4",  "chr5",  "chr6",  "chr7",  "chr8",  "chr9",  "chr10", "chr11", "chr12", "chr13", "chr14", "chr15", "chr16", "chr17", "chr18", "chr19", "chr20", "chr21", "chr22") )

@@

hs.CO.length_plot <- ggplot(Melted.Hs.male_MLH1, aes(y=CO, x=Chrm.id))+geom_jitter(aes( color=row.name, alpha = .2))+ 
  ggtitle("Human male CO Counts")+theme(legend.position="none")
#hs.CO.length_plot
```

@@

Heads of the cell-wide summary statistics.

@@

                            nbivs = length(unique(Chrm.id)),
                            total.SC = sum(axis.length),
                            mean.SC = mean(axis.length),
                            cV.SC = cv(axis.length)#this does within cell cV (dif than cov)
)
head(Hs.SC_cell_wide)

#the mus is rank based on SC, can I write in rank based on Chrm.id
for(cellys in as.character(unique(Melted.Hs.male_axis$row.name))){
  chrm.order = rank(Melted.Hs.male_axis$axis.length[whole.cell.data$row.name == cellys])
  Melted.Hs.male_axis$rank[Melted.Hs.male_axis$row.name == cellys] <- chrm.order  
}

#
two_groups2 <- transform(Melted.Hs.male_axis, category2 = factor(paste(row.name, Chrm.id)))
two_groups2 <- transform(two_groups2, category2 = reorder(category2, rank(desc(axis.length)), decreasing = FALSE))
#use desc for DECENDING order
MiniBiv_line.plot <-  ggplot(two_groups2, aes(category2, axis.length, group=row.name, color=row.name)) +  geom_line(alpha = 0.2) + 
  ggtitle("700 human male chrms ranked by Length")+ labs(x = "")+theme(axis.line=element_blank(),axis.text.x=element_blank(),
        axis.ticks=element_blank())

MiniBiv_line.plot

#order by chrm id (they are now an ordered factor)
#it might be kinda hard to subset the melted data, because they aren't organized into cells
#

Melted.Hs.male_axis.mini <- subset(Melted.Hs.male_axis, row.name < 101)

two_groups <- transform(Melted.Hs.male_axis.mini, category2 = factor(paste(row.name, Chrm.id)))
two_groups <- transform(two_groups, category2 = reorder(category2, rank(Chrm.id)))

MiniBiv_line.plot2 <-  ggplot(data=two_groups, aes(category2, axis.length, group=row.name, color=row.name)) +  geom_line(alpha = 0.5) +  ggtitle("700 human male chrms ranked by chrm ID")+ labs(x = "")+theme(axis.line=element_blank(),axis.text.x=element_blank(),
        axis.ticks=element_blank())

MiniBiv_line.plot2

#I could check a couple hundred, and remove cells with noticible spikes

@@

I added in single cell ranked axis curves. That is 22 axis/chromsomes for a human spermatocyte are ranked smallest to largest and I draw a curve over them. There are 700 cells, which obscure some patterns. A simple discription is that wider/thicker color indicates more variation in the final rank.

I'm trying to think more about the single cell looping factor. -- how would I discribe this, slope of ranked 
Sometimes the chrm id doesn't fit rank expectations.  (looking at the mini.biv plots, )

@@

I didn't test the axis length trait yet.  (See below for axis traits)

@@

```{r Mus.physcial.genome, echo=FALSE}
#distribution of genome sizes from reference Mus genome
#haploid genome size
refMus.geome_Mb_male <- c(195,182,160,152,150,145,129,125,131,122,120,120,125,104,91,95,91,61)# 1109.412

refMus.geome_Mb_XX <- c(195,182,160,152,150,145,129,125,131,122,120,120,125,104,91,95,91,61, 171)

#note, this ommits telomeres and centromeres!
#read the 
#on average 8Mb worth of centromeres and telomeres are added to each chrm
#i will try to find estimates of variation across indiviual chromosomes

Show a plot of the estimates chrm sizes. These values from MGI page omit centromere and telomere estimates. Read up on the T2T project lit to see if I can find better estimates of physical chormosome sizes.

Do I need most accurate estimate of Mb size variance? would I expect -- similar variance among cells? If i ran jack-knifing or sub-sampling permutations of the within cell covariance calculations -- if the covariation eatimate remained about the same, I would be more convinced that the 'packing' ratio is mostly uniform WITHIN cells.
Null/ comparison value might be, randomize cell.id for all bivalent observations.
(maybe I should first try this in human data? (minus the wacky looking cells / bad chrm id))

This might be the best thing to do to account for different physical sizes across other strains (non reference genome)

@@

                 totCO = sum(hand.foci.count),
                 total.SC = sum(SC.length),
                 mean_co = mean(hand.foci.count, na.rm = TRUE),
                 mean_sc = mean(SC.length),
                   var_co = var(hand.foci.count),
                   var_sc = var(SC.length),
                 cv_co.count = cv(hand.foci.count),
                 cv_SC.length = cv(SC.length)
)
#exclude last one
male.biv.count_table <- male.biv.count_table[!(is.na(male.biv.count_table$fileName) | male.biv.count_table$fileName==""), ]

male.biv.count_table$ratio.now <- var(refMus.geome_Mb_male) / male.biv.count_table$var_sc
male.biv.count_table$ratio.invert <-  male.biv.count_table$var_sc / var(refMus.geome_Mb_male)
#have to think of which is right

@@

                 totCO = sum(hand.foci.count, na.rm = TRUE),
                  total.SC = sum(SC.length, na.rm = TRUE),
                 mean_co = mean(hand.foci.count, na.rm = TRUE),
                 mean_sc = mean(SC.length, na.rm = TRUE),
                   var_co = var(hand.foci.count),
                   var_sc = var(SC.length),
                 cv_co.count = cv(hand.foci.count),
                 cv_SC.length = cv(SC.length)
)
#exclude last nas
female.biv.count_table <- female.biv.count_table[!(is.na(female.biv.count_table$fileName) | female.biv.count_table$fileName==""), ]

female.biv.count_table$ratio.now <- var(refMus.geome_Mb_XX) / female.biv.count_table$var_sc
female.biv.count_table$ratio.invert <-  female.biv.count_table$var_sc / var(refMus.geome_Mb_XX)

@@

Calculated the ratio of genome size variance (haploid chrm variance), (I think the right denominator is SC length variance)
most all are above 1 (for both female and male) -- indicates the packing ratio / packing process doesn't maintain (our guess about genome size of chromosomes)
